---
title: "p8105_hw3_jy3269"
author: "Jingyi Yao"
date: "`r Sys.Date()`"
output: github_document
---

```{r,warning=FALSE, message=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(p8105.datasets)

```

## Problem 2

### 1. import csv file
```{r}
data_1 <- read_csv("./data/accel_data.csv")
data_1

```

### 2. clean, tidy and wrangle data

```{r}
acc_1 <- data_1 %>% 
  janitor::clean_names() %>% 
  mutate(
    week = as.integer(week),
    day_id = as.integer(day_id),
    day = factor(day,levels= c("Monday", 
    "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday","Sunday")),
    weekday_vs_weekend = if_else(day == "Saturday" | day == "Sunday","weekend","weekday") # new variable indicate whether the day is weekend or weekday
  ) %>% 
  pivot_longer(                          # switch the data set from wider to longer
    activity_1:activity_1440,
    names_to = "minute",                 # activity_number in previous column names indicate the minute       
    names_prefix = "activity_",          # remove the activity_ prefix and preserve the number
    values_to = "activity_counts") %>%   # store the observed data in a new column names activity_counts
  mutate(minute = as.integer(minute))    # change data type

acc_1
```

Description:

*  structure of the data set:
  * the `acc_1` data set has  `r nrow(acc_1)`  rows (observations) and  `r ncol(acc_1)` columns (variables)
  * the variables in `acc_1` include :  `r names(acc_1)`
  * the `weekday_vs_weekend` variable is a new one added to the original data set, showing whether the day is a weekday or a weekend.
  * the original data set `data_1` has  `r ncol(data_1)`  variables. The last 1440 columns are the activity counts of each minute in a day. 
  * We used `pivot_longer` to transform the 1440 columns to store the data in a new variable named `activity_counts` and specify the minute each data belongs to in a new variable named `minute`.



### 3. calculate total activity each day + make a table

```{r}
acc_1 %>% 
  group_by(week,day) %>% 
  summarize(total_activity = sum(activity_counts)) %>% 
  arrange(week,day) %>%       # organize data in a tidy order
  pivot_wider(
    names_from = day,
    values_from = total_activity
  ) %>% 
  knitr::kable()
  
  
```


Description of the table :
*  The daily total activities tend to fluctuate within each week, thus there is no obvious trends.
*  The daily total activity counts on Saturday in Week 4 and 5 are both 1440 which means the activity counts is 1 for each minute. The data seems abnormal. I suppose that there might be something wrong with the Accelerator that makes it unable to collect data correctly.


To better discover the trends during each week, we may make a plot.


# 4. Plot the trend of each week
```{r}
acc_1 %>% 
  group_by(week,day) %>% 
  summarize(total_activity = sum(activity_counts)) %>% 
  arrange(week,day) %>%                             # use part of the code from making the previous table
  ggplot(aes(x = day, y = total_activity, color = week, group = week)) +
  geom_point() + geom_line() + facet_grid(~week) +  # place the 5 plots in a row in the order of weeks
  labs(
  title = "Activity Trend",
  x = "Day",
  y = "Daily Total Activity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
        legend.position = "none")


```


Description of the trend plot:
* In week 1 and 2, the total activity per day increases from Monday to Friday.
* In Week 3 and 4, the fluctuations are more irregular.
* Week 5 shows an approximately increasing trend from Monday to Friday.
* The abnormal data on the Saturday of Week 4 and 5 is more obvious in the plot.



### 5. plot daily activity
```{r}
acc_1 %>%  
  group_by(day_id) %>%
  ggplot(aes(x = minute,y = activity_counts,color = day)) +
  geom_line(alpha=.7) +
  theme(legend.position = "bottom")+
  labs(
    title = "Daily Activity",
    x = "Time in a Day",
    y = "Activity Count")+
  scale_x_continuous(
    breaks = c(0,180,360,540,720,900,1080,1260,1440),   # set the breaks of minutes in a day
    labels = c("12AM","3AM","6AM","9AM","12PM","3PM","6PM","9PM","12AM"),  # more readable x-axis labels
    limits = c(0,1440))
  
```


Description:
* Within each day, the man tend to be more active from 6 am to around 10 pm and less active from 11 pm to 6 am. This is probably because he goes to sleep at around 11 pm and wakes up t aound 6 am.
* There is a significant peak in pink around 11 am to 12 pm, which implies that the man is particularly active at noon on Sundays.
* Another peak in various colors is around 7 pm to 10 pm, which means that the man is very active almost each day during this time period. In the peak, there seems to be more blue lines than other colored lines, implying that the man is more active on Fridays during this time period.




## Problem 3

### 1. load `ny_noaa` data set and explore its structure
```{r}
data("ny_noaa")

num_weather_station <- length(unique(ny_noaa$id))  # number of missing values

no_missing = na.omit(ny_noaa)                      # data removed of the NAs
num_no_missing <- nrow(no_missing)                 # number of rows without missing values
missing_ proportion <- (nrow(ny_noaa) - nrow(no_missing))/nrow(ny_noaa)*100   # missing proportion


# missing proportion in each variables
ny_noaa %>% 
  summarise_at(vars(prcp:tmin), .funs = function(x) mean(is.na(x))) %>%
  round(digits = 3) %>% 
  knitr::kable()

```

Description of the raw dataset `ny_noaa`:
*  `ny_noaa` has  `r nrow(ny_noaa)` observations and  `r ncol(ny_noaa)` variables.
*  The variables are : names(ny_noaa)
  * id: Weather station ID
  * date: Date of observation
  * prcp: Precipitation (tenths of mm)
  * snow: Snowfall (mm)
  * snwd: Snow depth (mm)
  * tmax: Maximum temperature (tenths of degrees C)
  * tmin: Minimum temperature (tenths of degrees C)
*  The key variables are `prcp` `snow` `snwd` `tmax` and `tmin`
*  The observations ranging from 1981-01-01 to 2010-12-31
*  There are `r length(unique(ny_noaa$id))` different weather stations.


The missing value problem:
*  There are a total of `r nrow(ny_noaa)-nrow(no_missing)` rows that contains missing data, which is `r (nrow(ny_noaa)-nrow(no_missing))/nrow(ny_noaa)*100`% of the total observations. 
*  It means that more than half of the `ny_noaa` data set have missing values.Therefore, missing data is a big issue. 
*  The missing value proportion of the 5 key variables are shown in the table : 44% missing for max and min temperature , 23% for snow depth, 15% for snow, and 6% for precipitation.



### 2. clean  data
```{r}
noaa <- ny_noaa %>%
  janitor::clean_names() %>%
  separate(date, c("year", "month", "day"), sep = "-", convert = TRUE) %>%
  mutate(
    month = month.name[as.integer(month)],
    prcp = prcp / 10,
    tmax = as.numeric(tmax) / 10, 
    tmin = as.numeric(tmin) / 10    # switch to proper units
    )
noaa
```


*  `noaa` has `r nrow(noaa)` rows and `r ncol(noaa)` columns.
*  The variables in `noaa` are : `r names(noaa)`
*  the units of precipitation, min and max temperature  are adjusted.



### 3. snowfall
```{r}
noaa %>% 
  group_by(snow) %>% 
  summarise(frequency = n()) %>%  # count the occurrence of each observed snowfall data
  arrange(desc(frequency))        # arrange the data in descending order to find the most frequent ones

```

The most commonly observed values of snowfall variable in the data set is 0, which implies that there is no snow for most of the time in New York.


One reason for this is that winter only takes up a small proportion of the time range for the observations. And New York almost only snows in winter.


Another supplementary reason why New York does not snow very often even in winter is that the climate of New York state is humid continental. The average winter temperature is below 0 during January and February in most cities in the state. But it is usually several degrees above 0 along the Atlantic coastline, including New York City, making it hard to snow.




### 4. plot mean tmax

```{r}
noaa %>%
  filter(month == 'January' | month == 'July') %>%
  group_by(month, id, year) %>%
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = mean_tmax, color = month)) +
  geom_point(alpha = .2) +
  geom_smooth(se = F,color = "black")+    # add a black line to show the trend
  facet_grid(~month) +
  labs(x = "Year",
       y = "Average Max Temperature (C)",
       title = "Average Max Temperature in January and July") +
  theme_bw() +
  theme(legend.position = "bottom")

```
Description:

*  The mean tmax in January is much lower than that in July. 
*  The range of mean tmax in January is around -10 to 10 (degree C), in July is around 20 to 35 (degree C). Thus, Jan has a larger range than July.
*  There are outliers in both January and July. 
*  In January, there are extremely low mean tmax lower than -10 and extremely high mean tmax above 10. 
*  In July, there are also extremely high mean tmax close to 35 and extremely low mean tmax lower than 15.
*  The average mean tmax in January across the years is centered around 0 degree C, while that in July the average is centered around 27 degree C.
*  We can tell from the black line that the mean tax in January fluctuates more wildly than that in July.



## 5. plot tmax vs tmin +  distribution of snowfall

```{r}
tmax_tmin_plot <-                       # tmax vs tmin
  noaa %>% 
  ggplot(aes(x = tmax,y = tmin)) +
  geom_hex() +                          # use a hexagon plot
  geom_smooth(se = FALSE,na.rm = TRUE)
  labs(
    x = "tmax (degree C)",
    y = "tmin (degree C)",
    title = "tmax vs tmin ",
    caption = "Data from ny_noaa"
  )
  
snow_density_plot <-                    # snow density
  noaa %>% 
  filter(snow > 0 & snow < 100) %>% 
  mutate(year = as.factor(year)) %>% 
  ggplot(aes(x = snow, y = year,fill = year)) +
  geom_density_ridges() +
  labs(
    y = "Year",
    x = "Snowfall (mm)",
    title = "The distribution of snowfall",
    caption = "Data from ny_noaa") +
  theme_ridges() + theme(legend.position = "none")


tmax_tmin_plot + snow_density_plot


```
Description:
*  The left figure is a plot of tmax vs tmin from 1980 to 2010. It shows that most of the points fall into the center of the plot, where tmax ranges from 0 to 30 and tmin ranges from -10 to 15.
* The right figure is the distribution of snowfall over the years and each curve represents a year. 
* Although the distribution of snow are slightly different from year to year, they all share a similar pattern. 
* The pattern shows that : the highest density is between 0 mm to 25 mm, and the lowest is greater than 90 mm. It means that the snowfall is usually between 0 mm to 25 mm and rarely over 90 mm.





















